---
title: "HW1 Advanced R"
author: "Calvin Makelky, A01872013"
date: "October 4, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1
##a
```{r}
load(url("http://www.math.usu.edu/adele/IntroR/SFHousing.rda"))
sf = housing
names(sf)
sapply(sf, function(x) sum(is.na(x)))
```
The number of missing values for each variable is listed above.

#b
```{r}
tapply(sf$price, sf$county, median,  na.rm=TRUE)
```
The median housing price of each county is listed above.

#c
```{r}
citymean = tapply(sf$price, sf$city, mean,  na.rm=TRUE)
citysort = sort(citymean, decreasing=TRUE)
head(citysort, 10)
```
The top ten most expensive cities based on mean housing price is listed above.

#d
```{r}
zips = as.numeric(as.character(sf$zip))
sanfran = function(x){
    if ( abs(94118-x)>=17 | is.na(x)==TRUE )
        return(FALSE)

    else
        return(TRUE)
}
SFZip = sapply(zips, sanfran)
length(SFZip[SFZip==TRUE])
```
There are 8134 observations that fall into the SF area. 

#e
```{r}
tapply(sf$br, SFZip, mean)
```
On average there is about .7 less bedrooms in houses in San Fransisco than in other cities outside of San Fransisco.

#2
#a
```{r}
gdpdata= read.table("gdp.txt")
athdata= read.table("athletes2016.txt")
popdata= read.table("population.txt")
```
Gdp is a discrete quantitative variable.
Population is a discrete quantitative variable.
Total is a discrete quantitative variable.

#b
```{r}
colnames(gdpdata) = c("country", "gdp")
athdata = athdata[,-(2:5)]
colnames(athdata) = c("country", "total")
colnames(popdata) = c("country", "population")
plot(popdata$population, gdpdata$gdp)
```
This plot violates all three properties of a good graph.
1) The data does not stand out, as it is in black and white.
2) The plot does not facilitate comparison because of the lack of color and more importantly how 3 super large or rich countries obscure the bulk of the data
3) The plot is not information rich at all. It lacks a title and a clear x and y axis title. There's also no caption to decribe what is going on. This is the gdp and population of what countries?

#c
```{r}
gdpPP = gdpdata$gdp / popdata$population
library(RColorBrewer) 
mycolors = brewer.pal(9, "Set3")
symbols(log(popdata$population), log(gdpPP), circles=athdata$total, 
        col=mycolors, bg=mycolors, inches=.6 )
```

#d
```{r}
symbols(log(popdata$population), log(gdpPP), inches =.6, 
circles=athdata$total, bg=mycolors, main="GDP Per Person vs Population", 
xlab="Log of Population of Country", ylab="Log of GDP Per Person in Country")

countries5 = c(198, 41, 87, 111, 124)

text(log(popdata[countries5,2]), log(gdpPP[countries5]), 
     popdata[countries5,1], cex=1, pos=3, col="red")

```
Seems to be a relationship between number of medals and population, while none between GDP per person and number of medals won.
